---
title: "Data_wrangling_zoop_fish"
author: "Denise-Colombano"
date: "05/03/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load libraries
```{r}
library(tidyverse)
library(ggridges)

library(zooper)
library(LTMRdata)

# update Zooper package
#devtools::install_github("InteragencyEcologicalProgram/zooper")

# update LTMRdata
#devtools::install_github("sbashevkin/LTMRdata")


library(DT)
library(leaflet)
library(sf)
library(janitor)

library(tidygraph)
library(sfnetworks)
```

# Import data from Zooper
```{r}
MyZoops <- Zoopsynther(Data_type = "Community", 
                       Sources = "EMP")
View(MyZoops)

library(dplyr)
library(ggplot2)
library(RColorBrewer)
set.seed(16)

# MyZoops%>%
#   filter(!is.na(SalSurf))%>%
#   mutate(Salinity_zone=case_when(
#     SalSurf < 0.5 ~ "Freshwater",
#     SalSurf > 0.5 & SalSurf < 6 ~ "Low salinity zone",
#     SalSurf > 6 ~ "High salinity zone"
#   ))%>%
#   mutate(Salinity_zone=factor(Salinity_zone, levels=c("Freshwater", "Low salinity zone", "High salinity zone")))%>%
#   group_by(Year,Phylum, Class, Order, Family, Genus, Species, Lifestage, Taxlifestage, Salinity_zone)%>%
#   summarise(CPUE=mean(CPUE, na.rm=T))%>%
#   ungroup()%>%
#   arrange(Phylum, Class, Order, Family, Genus, Species, Lifestage)%>%
#   mutate(Taxlifestage=factor(Taxlifestage, unique(Taxlifestage)))%>%
#   ggplot(aes(x=Year, y=CPUE))+
#   geom_bar(stat="identity", color="white", size=0.01, aes(fill=Taxlifestage))+
#   facet_wrap(~Salinity_zone, nrow=1)+
#   coord_cartesian(expand=0)+
#   scale_x_continuous(breaks = function(x) unique(floor(pretty(seq(min(x), max(x)), n=3))), expand=c(0,0))+
#   scale_fill_manual(values=sample(colorRampPalette(brewer.pal(12, "Set3"))(length(unique(MyZoops$Taxlifestage)))), 
#                     name="Taxa and life stage", 
#                     guide = guide_legend(ncol=3, title.position = "top", title.hjust = 0.5))+
#   ylab(bquote(Average~CPUE~"("*Catch*"/"*m^3*")"))+
#   theme_bw()+
#   theme(panel.grid=element_blank(), text=element_text(size=14), legend.text = element_text(size=8), 
#         legend.key.size = unit(10, "points"), strip.background=element_blank(), legend.position = "bottom", 
#         legend.background = element_rect(color="black"), axis.text.x=element_text(angle=45, hjust=1))


MyZoopsOrders <- MyZoops %>% 
  distinct(Order)
MyZoopsPhylum <- MyZoops %>% 
  distinct(Phylum)
```

# Filter zoops
```{r}
# Filter out taxa of interest - using the mesozooplankton net
MyZoopsT <- MyZoops %>% 
  filter(
    Order == "Calanoida" |
      Order == "Cyclopoida" | 
      Order == "Cladocera" | 
      Order == "Mysida" |
      Phylum == "Rotifera"
  )  %>% 
  filter(!is.na(SalSurf))%>%
  mutate(Salinity_zone=case_when(
    SalSurf < 0.5 ~ "Freshwater",
    SalSurf > 0.5 & SalSurf < 6 ~ "Low salinity zone",
    SalSurf > 6 ~ "High salinity zone"
  ))%>%
  mutate(Salinity_zone=factor(Salinity_zone, levels=c("Freshwater", 
                                                      "Low salinity zone", 
                                                      "High salinity zone"))) %>% 
  mutate(Period= case_when(
    Year < 1987 ~ "Period1",
    Year >= 1987 & Year <= 1994 ~ "Period2",
    Year >= 1995 & Year <= 2006 ~ "Period3",
    Year >= 2007 & Year <= 2016 ~ "Period4",
    Year > 2016 ~ "Period5"
  )) %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date), Month=lubridate::month(Date))
head(MyZoopsT)
dim(MyZoopsT)

# Filter out core stations
zoop_month <- read_csv("data/stations/stations_zoop_month.csv") %>% 
  mutate(Category="Zoop", Survey="EMP Zoop")

# Match them - only core stations (n=15)
MyZoopsF <- MyZoopsT %>% 
  right_join(zoop_month)
 
head(MyZoopsF)
dim(MyZoopsF)

MyZoopsMeso <- MyZoopsF %>% 
  filter(SizeClass == "Meso")

MyZoopsMacro <- MyZoopsF %>% 
  filter(SizeClass == "Macro")
```

# Map
```{r}
library(DT)
library(leaflet)
library(sf)
library(janitor)

# 15 core stations
MyZoopsMap <- MyZoopsF %>% 
  distinct(Source, Station, Latitude, Longitude) %>% 
  arrange(desc(Longitude))

leaflet() %>% 
  addTiles() %>% 
    addCircleMarkers(data = MyZoopsMap, 
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = 4,
                         weight = 0.1,
                          color= "black",
                          fillOpacity = 0.7)
```

## sampling density
```{r}
# Which months were consistently sampled over time?
# Sampling <- MyZoopsF %>% 
#   mutate(Month=lubridate::month(Date), Year=lubridate::year(Date)) %>% 
#   group_by(Station, Month, Year) %>% 
#   summarize(totCPUE=sum(CPUE)) %>%
#   group_by(Station, Year) %>% 
#   summarize(Months=tally(Month))

Sampling <- MyZoopsF %>% 
  select(Source, SizeClass, Station, Date, Julian, Month) %>% 
  distinct() %>% 
  mutate(Year=lubridate::year(Date)) %>% 
  filter(!is.na(Year)) %>% 
  filter(SizeClass=="Meso")

ggplot(Sampling, aes(Julian))+
  geom_histogram(binwidth =1, aes(fill=Station))+
  facet_wrap(Year~.)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  labs(title="Zoop", subtitle="Meso")

Sampling_b <- MyZoopsF %>% 
  select(Source, SizeClass, Station, Date, Julian, Month) %>% 
  distinct() %>% 
  mutate(Year=lubridate::year(Date)) %>% 
  filter(!is.na(Year)) %>% 
  filter(SizeClass=="Meso", Year==2015)

ggplot(Sampling_b, aes(Julian))+
  geom_histogram(binwidth =1, aes(fill=Station))+
  #facet_wrap(~.)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  labs(title="Zoop 2015", subtitle="Meso")
```


# Groups

## calanoida
```{r}
Calanoida <- MyZoopsMeso %>% 
  filter(Order=="Calanoida") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date))

ggplot(Calanoida, aes(Julian, CPUE))+
  geom_point(aes(color=Taxlifestage))+
  facet_grid(Year~Taxlifestage)

Calanoida_genus <- Calanoida %>% 
  filter(Genus=="Pseudodiaptomus" |
           Genus=="Eurytemora" | 
           Genus == "Sinocalanus" | 
           Genus == "Acartiella") %>% 
  group_by(Genus, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Calanoida_genus, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(Genus~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Calanoida", x="Julian date")
```

## cyclopoida
```{r}
Cyclopoida <- MyZoopsMeso %>% 
  filter(Order=="Cyclopoida") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date))

ggplot(Cyclopoida, aes(Julian, CPUE))+
  geom_point(aes(color=Taxlifestage))+
  facet_grid(Year~Taxlifestage)

Cyclopoida_genus <- Cyclopoida %>% 
  filter(Genus=="Acanthocyclops" |
           Genus=="Limnoithona" |
           Genus =="Oithona" ) %>% 
  group_by(Genus, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Cyclopoida_genus, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(Genus~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Cyclopoida", x="Julian date")
```


## cladocera
```{r}
Cladocera <- MyZoopsMeso %>% 
  filter(Order=="Cladocera") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date))

ggplot(Cladocera, aes(Julian, CPUE))+
  geom_point(aes(color=Taxlifestage))+
  facet_grid(Year~Taxlifestage)

Cladocera_genus <- Cladocera %>% 
  filter(Species == "Bosmina longirostris" | # only species IDed
           Taxname == "Daphnia_UnID") %>% #Daphnia
  group_by(Taxname, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE)) %>% 
  mutate(Taxname2= ifelse(Taxname=="Daphnia_UnID", "Daphnia spp.", Taxname))

ggplot(Cladocera_genus, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(Taxname2~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Cladocera", x="Julian date")
```


## mysida
```{r}
Mysida <- MyZoopsMacro %>% 
  filter(Order=="Mysida") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date))

ggplot(Mysida, aes(Julian, CPUE))+
  geom_point(aes(color=Taxlifestage))+
  facet_grid(Year~Taxlifestage)

Mysida_genus <- Mysida %>% 
  filter(Genus == "Alienacanthomysis" |
           Genus == "Hyperacanthomysis" |
           Genus == "Neomysis" |
           Genus == "Orientomysis") %>% 
  group_by(Genus, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Mysida_genus, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(Genus~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Mysida", x="Julian date")
```


## rotifera
```{r}
Rotifera <- MyZoopsF %>% 
  filter(Phylum=="Rotifera") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date))

ggplot(Rotifera, aes(Julian, CPUE))+
  geom_point(aes(color=Taxlifestage))+
  facet_grid(Year~Taxlifestage)

Rotifera_genus <- Rotifera %>% 
  group_by(Taxlifestage, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Rotifera_genus, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(.~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Rotifera", x="Julian date")
```


# Species

## Mesozooplankton

### Calanoid and cyclopoids
```{r}
# Preliminary analysis - by drought vs. non-drought period
# Filter species of interest
MyZoopsMeso_SppMonth <- MyZoopsMeso %>% 
  filter(Species == "Eurytemora affinis" |
          Species == "Pseudodiaptomus forbesi" |
           Species=="Sinocalanus doerrii" |
           Species=="Limnoithona tetraspina" |
           Species=="Limnoithona tetraspina") %>%
  group_by(Species, Month, Period,Year) %>% #, Period
  summarize(meanCPUE=mean(CPUE, na.rm=T))

# Identify the peaks for each species x period combination
SppMonth_peaks <- MyZoopsMeso_SppMonth %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Species, Period,Month) %>% #Period, 
  summarize(meanCPUE=median(meanCPUE)) %>% # median
  filter(meanCPUE== max(meanCPUE)) %>% 
  filter(meanCPUE != 0)

# Plot monthly trends for all years
ggplot(MyZoopsMeso_SppMonth, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_vline(data=SppMonth_peaks, aes(xintercept = Month), lty=2, col="red")+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Species~Period, scales = "free_y")+ #Period
  xlim(2.5,10.5)+
  scale_fill_viridis_c()+
  labs(title="Common mesozooplankton species")


# Identify the peaks but with standardized CPUE values for each species
SppMonth_z <- MyZoopsMeso_SppMonth %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Species,Period,Month) %>% 
  summarize(meanCPUE=mean(meanCPUE)) %>%
  group_by(Species, Period) %>% 
  mutate(meanCPUE2= mean(meanCPUE, na.rm=T), sd=sd(meanCPUE, na.rm = T)) %>% 
  mutate(CPUEz= (meanCPUE - meanCPUE2) / sd) %>% 
  ungroup()

SppMonth_peaks_z <- SppMonth_z %>% 
  group_by(Species, Period) %>% 
  filter(CPUEz== max(CPUEz)) %>% 
  filter(CPUEz != 0)

# zscore CPUE
ggplot(SppMonth_z, aes(x=Month, y=CPUEz, height=CPUEz, group=Period, fill=Period))+
  geom_hline(aes(yintercept=0), lty=2, col="gray50")+
  geom_ridgeline(min_height = -5, stat="identity")+
  facet_grid(.~Species)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)

# mean CPUE
ggplot(SppMonth_z, aes(x=Month, y=meanCPUE, height=meanCPUE, group=Period, fill=Period))+
  geom_ridgeline(min_height = -2, stat="identity")+
  facet_grid(.~Species)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)

###
# Next analysis - Identify peak timing for each year
```

### Cladocerans
```{r}
# Preliminary analysis - by drought vs. non-drought period
# Filter species of interest
MyZoopsMeso_SppMonth2 <- MyZoopsMeso %>% 
  filter(Taxname == "Bosmina longirostris" |
           Taxname == "Daphnia_UnID") %>%
  group_by(Taxname, Month, Year, Period) %>% 
  summarize(meanCPUE=mean(CPUE, na.rm=T)) %>% 
  mutate(Taxname=ifelse(Taxname=="Daphnia_UnID", "Daphnia spp.", Taxname))

# Identify the peaks for each species x period combination
SppMonth_peaks2 <- MyZoopsMeso_SppMonth2 %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Taxname, Period, Month) %>% 
  summarize(meanCPUE=median(meanCPUE)) %>% # median
  filter(meanCPUE== max(meanCPUE)) %>% 
  filter(meanCPUE != 0)

# Plot monthly trends for each period
ggplot(MyZoopsMeso_SppMonth2, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_vline(data=SppMonth_peaks2, aes(xintercept = Month), lty=2, col="red")+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Taxname~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  scale_fill_viridis_c()+
  labs(title="Common cladocerans")


# Identify the peaks but with standardized CPUE values for each species
SppMonth_z2 <- MyZoopsMeso_SppMonth2 %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Taxname,Period,Month) %>% 
  summarize(meanCPUE=mean(meanCPUE)) %>%
  group_by(Taxname, Period) %>% 
  mutate(meanCPUE2= mean(meanCPUE, na.rm=T), sd=sd(meanCPUE, na.rm = T)) %>% 
  mutate(CPUEz= (meanCPUE - meanCPUE2) / sd) %>% 
  ungroup()

SppMonth_peaks_z2 <- SppMonth_z2 %>% 
  group_by(Taxname, Period) %>% 
  filter(CPUEz== max(CPUEz)) %>% 
  filter(CPUEz != 0)

# zscore CPUE
ggplot(SppMonth_z2, aes(x=Month, y=CPUEz, height=CPUEz, group=Period, fill=Period))+
  geom_hline(aes(yintercept=0), lty=2, col="gray50")+
  geom_ridgeline(min_height = -5, stat="identity")+
  facet_grid(.~Taxname)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)

# mean CPUE
ggplot(SppMonth_z2, aes(x=Month, y=meanCPUE, height=meanCPUE, group=Period, fill=Period))+
  geom_ridgeline(min_height = -2, stat="identity")+
  facet_grid(.~Taxname)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)


###
# Next analysis - Identify peak timing for each year
```


## Macrozooplankton

```{r}
# Preliminary analysis - by drought vs. non-drought period
# Filter species of interest
MyZoopsMacro_SppMonth <- MyZoopsMacro %>% 
  filter(Taxname == "Hyperacanthomysis longirostris" |
           Taxname == "Neomysis kadiakensis" |
           Taxname == "Neomysis mercedis") %>%
  group_by(Taxname, Month, Year, Period) %>% 
  summarize(meanCPUE=mean(CPUE, na.rm=T))

# Identify the peaks for each species x period combination
SppMonthMacro_peaks <- MyZoopsMacro_SppMonth %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Taxname, Period, Month) %>% 
  summarize(meanCPUE=median(meanCPUE)) %>% # median 
  filter(meanCPUE== max(meanCPUE)) %>% 
  filter(meanCPUE != 0)

# Plot monthly trends for each period
ggplot(MyZoopsMacro_SppMonth, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_vline(data=SppMonthMacro_peaks, aes(xintercept = Month), lty=2, col="red")+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Taxname~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  scale_fill_viridis_c()+
  labs(title="Common macrozooplankton species")


# Identify the peaks but with standardized CPUE values for each species
SppMonthMacro_z <- MyZoopsMacro_SppMonth %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Taxname,Period,Month) %>% 
  summarize(meanCPUE=mean(meanCPUE)) %>% 
  group_by(Taxname, Period) %>% 
  mutate(meanCPUE2= mean(meanCPUE, na.rm=T), sd=sd(meanCPUE, na.rm = T)) %>% 
  mutate(CPUEz= (meanCPUE - meanCPUE2) / sd) %>% 
  ungroup()

SppMonth_peaks_z <- SppMonth_z %>% 
  group_by(Taxname, Period) %>% 
  filter(CPUEz== max(CPUEz)) %>% 
  filter(CPUEz != 0)

# zscore CPUE
ggplot(SppMonthMacro_z, aes(x=Month, y=CPUEz, height=CPUEz, group=Period, fill=Period))+
  geom_hline(aes(yintercept=0), lty=2, col="gray50")+
  geom_ridgeline(min_height = -5, stat="identity")+
  facet_grid(.~Taxname)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)

# mean CPUE
ggplot(SppMonthMacro_z, aes(x=Month, y=meanCPUE, height=meanCPUE, group=Period, fill=Period))+
  geom_ridgeline(min_height = -2, stat="identity")+
  facet_grid(.~Taxname)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)


###
# Next analysis - Identify peak timing for each year
```

# Species

## Eury
```{r}
Eury <- MyZoopsMeso %>% 
  filter(Species=="Eurytemora affinis") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date), Month=lubridate::month(Date))

Eury_mean <- Eury %>% 
  group_by(Species, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Eury_mean, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(.~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Eurytemora affinis", x="Julian date")

Eury_month <- Eury %>% 
  group_by(Species, Month, Year, Period, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

# identify peak month in each period - TBD
#Eury_peaks <- Eury_month %>% 
 # group_by(Period)

# monthly
ggplot(Eury_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(.~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Eurytemora affinis")+
  scale_fill_viridis_c()

# with salinity
ggplot(Eury_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Salinity_zone~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Eurytemora affinis")+
  scale_fill_viridis_c()
```

## Pseudo
```{r}
Pseudo <- MyZoopsF %>% 
  filter(Species=="Pseudodiaptomus forbesi") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date), Month=lubridate::month(Date))

Pseudo_mean <- Pseudo %>% 
  group_by(Species, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Pseudo_mean, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(.~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Pseudodiaptomus forbesi", x="Julian date")

Pseudo_month <- Pseudo %>% 
  group_by(Species, Month, Year, Period, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

# identify peak month in each period - TBD

# monthly
ggplot(Pseudo_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(.~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Pseudodiaptomus forbesi")+
  scale_fill_viridis_c()

# with salinity
ggplot(Pseudo_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Salinity_zone~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Pseudodiaptomus forbesi")+
  scale_fill_viridis_c()
```

## Sino
```{r}
Sino <- MyZoopsF %>% 
  filter(Species=="Sinocalanus doerrii") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date), Month=lubridate::month(Date))

Sino_mean <- Sino %>% 
  group_by(Species, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Sino_mean, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(.~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Sinocalanus doerrii", x="Julian date")

Sino_month <- Sino %>% 
  group_by(Species, Month, Year, Period, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

# identify peak month in each period - TBD

# monthly
ggplot(Sino_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(.~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Sinocalanus doerrii")+
  scale_fill_viridis_c()

# with salinity
ggplot(Sino_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Salinity_zone~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Sinocalanus doerrii")+
  scale_fill_viridis_c()
```


## Limno
```{r}
Limno <- MyZoopsF %>% 
  filter(Species=="Limnoithona tetraspina") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date), Month=lubridate::month(Date))

Limno_mean <- Limno %>% 
  group_by(Species, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Limno_mean, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(.~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Limnoithona tetraspina", x="Julian date")

Limno_month <- Limno %>% 
  group_by(Species, Month, Year, Period, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

# identify peak month in each period - TBD

# monthly
ggplot(Limno_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(.~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Limnoithona tetraspina")+
  scale_fill_viridis_c()

# with salinity
ggplot(Limno_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Salinity_zone~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Limnoithona tetraspina")+
  scale_fill_viridis_c()
```




## Bosmina
```{r}
Bosmina <- MyZoopsF %>% 
  filter(Species=="Bosmina longirostris") %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date), Month=lubridate::month(Date))

Bosmina_mean <- Bosmina %>% 
  group_by(Species, Julian, Year, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

ggplot(Bosmina_mean, aes(Julian, meanCPUE, group=Year))+
  geom_line(aes(color=Year), alpha=0.7)+
  facet_grid(.~Salinity_zone, scales = "free_y")+
  scale_color_viridis_c()+
  labs(title="Bosmina longirostris", x="Julian date")


Bosmina_month <- Bosmina %>% 
  group_by(Species, Month, Year, Period, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE))

# identify peak month in each period - TBD

# monthly
ggplot(Bosmina_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(.~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Bosmina longirostris")+
  scale_fill_viridis_c()

# with salinity
ggplot(Bosmina_month, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Salinity_zone~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  labs(title="Bosmina longirostris")+
  scale_fill_viridis_c()
```


# Import data from LTMR
```{r}
library(LTMRdata)

BayStudyData <- fish(sources=c("Baystudy"),
             species=NULL,
             convert_lengths=TRUE,
             size_cutoff=40,
             zero_fill = TRUE,
             remove_unknown_lengths = TRUE,
             univariate = FALSE,
             quiet = FALSE)
str(BayStudyData)
```

# Filter fishes
```{r}
# Filter out forage fish species of interest
BayStudySpp <- BayStudyData %>% 
  mutate(Year=lubridate::year(Date), Julian=lubridate::yday(Date), Month=lubridate::month(Date)) %>% 
  filter(Method=="Midwater trawl") %>% 
  filter(Taxa == "Spirinchus thaleichthys" | 
           Taxa == "Hypomesus transpacificus" |
           Taxa == "Morone saxatilis" |
           Taxa == "Alosa sapidissima" |
           Taxa == "Dorosoma petenense" |
           Taxa == "Clupea pallasii" |
           Taxa == "Engraulis mordax") %>% 
    filter(!is.na(Sal_surf))%>%
  mutate(Salinity_zone=case_when(
    Sal_surf < 0.5 ~ "Freshwater",
    Sal_surf > 0.5 & Sal_surf < 6 ~ "Low salinity zone",
    Sal_surf > 6 ~ "High salinity zone"
  ))%>%
  mutate(Salinity_zone=factor(Salinity_zone, levels=c("Freshwater", 
                                                      "Low salinity zone", 
                                                      "High salinity zone"))) %>% 
  mutate(Period= case_when(
    Year < 1987 ~ "Period1",
    Year >= 1987 & Year <= 1994 ~ "Period2",
    Year >= 1995 & Year <= 2006 ~ "Period3",
    Year >= 2007 & Year <= 2016 ~ "Period4",
    Year > 2016 ~ "Period5"
  )) 
head(BayStudySpp)

# import Bay Study core station list - 19 total
baystudy_annual <- read_csv("data/stations/stations_fish_BayStudy_annual.csv") %>% 
  filter(Station>300) %>%  # San Pablo Bay lowest downstream extent
  mutate(Station=as.character(Station)) %>% 
  select(Station, Latitude, Longitude)
dim(baystudy_annual)

# Join them together- filter out catches from 19 core stations
BayStudyF <- BayStudySpp %>% 
  right_join(baystudy_annual)
head(BayStudyCore)
```


# Map
```{r}
library(DT)
library(leaflet)
library(sf)
library(s2)
library(janitor)
library(lwgeom)

# 15 core stations
BayStudyMap <- BayStudyF %>% 
  distinct(Source, Station, Latitude, Longitude)

leaflet() %>% 
  addTiles() %>% 
    addCircleMarkers(data = BayStudyMap, 
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = 4,
                         weight = 0.1,
                          color= "black",
                          fillOpacity = 0.7)
```

## sampling
```{r}
Sampling2 <- BayStudyF %>% 
  select(Source, Method, Station, Date, Julian, Year) %>% 
  distinct() %>%  
  filter(!is.na(Year)) %>% 
  filter(Method=="Midwater trawl")

ggplot(Sampling2, aes(Julian))+
  geom_histogram(binwidth =1, aes(fill=Station))+
  facet_wrap(Year~.)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  labs(title="Bay study", subtitle="Midwater trawl")


Sampling2_b <- BayStudyF %>% 
  select(Source, Method, Station, Date, Julian, Year) %>% 
  distinct() %>%  
  filter(!is.na(Year)) %>% 
  filter(Method=="Midwater trawl", Year=="2015")

ggplot(Sampling2_b, aes(Julian))+
  geom_histogram(binwidth =1, aes(fill=Station))+
  #facet_wrap(Year~.)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  labs(title="Bay study 2015", subtitle="Midwater trawl")
```



# Age class cutoffs
From Kathy Hieb at CDFW


Example - Longfin Smelt for June - All lengths are FL
Age-0 includes min length to  < age-0 cutoff				    40-66
Age 1+ >= age-cutoff					                          >=67
Age1 >= age-0 to <age-1 cutoff					               67-107
Age2+ >= age-1 cutoff					                          >=108

```{r}
BayStudyAges <- read_csv("BayStudy_AgeCutOffs.csv") %>% 
  select(AlphaCode:Taxa) %>% 
  rename(Month=Survey) # Survey num = month
View(BayStudyAges)

BayStudyFA <- BayStudyF %>% 
  mutate(AgeClass= case_when(
    # American Shad Age-0 vs. Age-1+
    Taxa == "Alosa sapidissima" & Month <=4 & Length <= 30 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 5 & Length <= 60 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 6 & Length <= 70 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 7 & Length <= 95 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 8 & Length <= 110 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 9 & Length <= 145 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 10 & Length <= 165 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 11 & Length <= 170 ~ "Age0",
    Taxa == "Alosa sapidissima" & Month == 12 & Length <= 170 ~ "Age0",
    # Delta Smelt Age-0 vs. Age-1+
    Taxa == "Hypomesus transpacificus" & Month == 1 & Length <= 40 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 2 & Length <= 50 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 3 & Length <= 50 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 4 & Length <= 50 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 5 & Length <= 50 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 6 & Length <= 55 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 7 & Length <= 65 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 8 & Length <= 70 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 9 & Length <= 75 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 10 & Length <= 80 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 11 & Length <= 80 ~ "Age0",
    Taxa == "Hypomesus transpacificus" & Month == 12 & Length <= 80 ~ "Age0",
    # Longfin Smelt Age-0 vs. Age-1+
    Taxa == "Spirinchus thaleichthys" & Month == 1 & Length <= 40 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 2 & Length <= 42 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 3 & Length <= 46 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 4 & Length <= 52 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 5 & Length <= 59 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 6 & Length <= 67 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 7 & Length <= 71 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 8 & Length <= 75 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 9 & Length <= 80 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 10 & Length <= 83 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 11 & Length <= 85 ~ "Age0",
    Taxa == "Spirinchus thaleichthys" & Month == 12 & Length <= 87 ~ "Age0",
    # Northern Anchovy Age-0 vs. Age-1+
    Taxa == "Engraulis mordax" & Length <= 90 ~ "Age0",
    # Pacific Herring Age-0 vs. Age-1+ 
    Taxa == "Clupea pallasii" & Month == 1 & Length <= 45 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 2 & Length <= 50 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 3 & Length <= 70 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 4 & Length <= 85 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 5 & Length <= 95 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 6 & Length <= 105 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 7 & Length <= 110 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 8 & Length <= 120 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 9 & Length <= 125 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 10 & Length <= 125 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 11 & Length <= 135 ~ "Age0",
    Taxa == "Clupea pallasii" & Month == 12 & Length <= 140 ~ "Age0",
    # Striped Bass 
    Taxa == "Morone saxatilis" & Month == 1 & Length <= 165 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 2 & Length <= 165 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 3 & Length <= 170 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 4 & Length <= 175 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 5 & Length <= 60 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 6 & Length <= 80 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 7 & Length <= 90 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 8 & Length <= 125 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 9 & Length <= 135 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 10 & Length <= 145 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 11 & Length <= 155 ~ "Age0",
    Taxa == "Morone saxatilis" & Month == 12 & Length <= 165 ~ "Age0",
    # Threadfin Shad
    Taxa == "Dorosoma petenense" & Length <= 500 ~ "Age0"
  )) %>% 
  filter(AgeClass=="Age0")
```




# Age-0 Species
```{r}
# Filter species of interest
Fish_SppMonth <- BayStudyFA %>% 
  group_by(Taxa, Month, Year, Period) %>% 
  summarize(meanCPUE=mean(Count, na.rm=T)) %>% 
  filter(Taxa>0) # remove NA

# Identify the peaks for each species x period combination
Fish_SppMonth_peaks <- Fish_SppMonth %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Taxa, Period, Month) %>% 
  summarize(meanCPUE=median(meanCPUE)) %>% # median
  filter(meanCPUE== max(meanCPUE)) %>% 
  filter(meanCPUE != 0)

# Plot monthly trends for each period
ggplot(Fish_SppMonth, aes(Month, meanCPUE, group=Month, fill=Month))+
  geom_vline(data=Fish_SppMonth_peaks, aes(xintercept = Month), lty=2, col="red")+
  geom_boxplot(size=0.3, outlier.shape = 21, outlier.size = 0.5)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  facet_grid(Taxa~Period, scales = "free_y")+
  xlim(2.5,10.5)+
  scale_fill_viridis_c()+
  labs(title="Common forage fish species")


# Identify the peaks but with standardized CPUE values for each species
Fish_SppMonth_z <- Fish_SppMonth %>% 
  ungroup() %>%
  filter(Month>2 & Month < 11) %>% 
  group_by(Taxa,Period,Month) %>% 
  summarize(meanCPUE=mean(meanCPUE)) %>%
  group_by(Taxa, Period) %>% 
  mutate(meanCPUE2= mean(meanCPUE, na.rm=T), sd=sd(meanCPUE, na.rm = T)) %>% 
  mutate(CPUEz= (meanCPUE - meanCPUE2) / sd) %>% 
  ungroup()

Fish_SppMonth_peaks_z <- Fish_SppMonth_z %>% 
  group_by(Taxa, Period) %>% 
  filter(CPUEz== max(CPUEz)) %>% 
  filter(CPUEz != 0)

# zscore CPUE
ggplot(Fish_SppMonth_z, aes(x=Month, y=CPUEz, height=CPUEz, group=Period, fill=Period))+
  geom_hline(aes(yintercept=0), lty=2, col="gray50")+
  geom_ridgeline(min_height = -10, stat="identity")+
  facet_grid(.~Taxa)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)

# mean CPUE
ggplot(Fish_SppMonth_z, aes(x=Month, y=meanCPUE, height=meanCPUE, group=Period, fill=Period))+
  geom_ridgeline(min_height = -10, stat="identity")+
  facet_grid(.~Taxa)+
  theme_bw()+
  scale_fill_viridis_d(alpha=0.5)
```





```{r}
# Longfin smelt
BayStudySuisun <- fish(sources=c("Suisun", "Baystudy"),
                       species="Spirinchus thaleichthys",
             convert_lengths=TRUE,
             remove_unconverted_lengths=FALSE,
             size_cutoff=40,
             zero_fill = TRUE,
             remove_unknown_lengths = TRUE,
             univariate = FALSE,
             quiet = FALSE)
BayStudySuisun

# Longfin smelt
Lonsme <- BayStudySuisun %>% 
  filter(Method=="Otter trawl")
Lonsme$Month <- lubridate::month(Lonsme$Date)
Lonsme$Year <- lubridate::year(Lonsme$Date)

Lonsme2 <- Lonsme %>% 
  filter(Month>3 & Month <11) %>%   # most consistent trawling effort for Bay Study
  group_by(Year, Source, Station) %>% 
  summarize(Count=sum(Count), Tow_area=sum(Tow_area), Tow_duration=sum(Tow_duration)) %>% 
  mutate(CPUE= ifelse(Source=="Bay Study", Count/Tow_area, Count/Tow_duration)) %>%
  mutate(Species="Longfin Smelt", Method="Otter trawl") %>% 
  filter(Year>1979) 
View(Lonsme2)

ggplot(Lonsme2, aes(Year, CPUE))+
  geom_line(aes(color=Station)) + 
  theme_bw()
```



# Zooplankton varpart

Step 1: calculate distances within the deltamapr polygon
```{r}
library(deltamapr)
require(ggplot2)
require(sf)
```

## spatial extent
```{r}
# Extract subregions of interest and combine
regions <- deltamapr::R_EDSM_Subregions_Mahardja
st_crs(regions) <- 26910
plot(regions)

waterways <- deltamapr::WW_Delta 
st_transform(waterways, crs=4326)

waterways2 <- waterways %>% 
  mutate(Extent="Waterway") %>% 
  group_by(Extent) %>% 
  summarise()%>% 
  st_make_valid()

ggplot(regions)+
  geom_sf(aes(fill=Region))+
  geom_sf_label(aes(label=SubRegion), size=2) +
  theme_bw()+
  ggthemes::scale_fill_colorblind()

Regions_mismatches <- regions %>%
  filter(SubRegion %in% c("San Pablo Bay", "Carquinez Strait", "West Suisun Bay", 
                          "Mid Suisun Bay", "Honker Bay", "Confluence", 
                          "Lower Sacramento River", "Lower San Joaquin River", 
                          "Sacramento River near Rio Vista", "San Joaquin River at Twitchell Island",
                          "Grizzly Bay")) %>% 
  mutate(Extent="Mismatch Study Extent") %>% 
  group_by(Extent) %>% 
  summarise()%>% 
  st_make_valid() %>% 
  st_transform(crs=4326) %>% 
  sf::st_as_s2(rebuild=TRUE) %>% 
  sf::st_as_sfc()
plot(Regions_mismatches)


# clip the waterway polygon to the region polygon
#ww_clip <- st_intersection(Regions_mismatches, waterways2)
#plot(ww_clip)
```

## 1 - zoop distances
```{r}
station_zoop <- tibble(lat=MyZoopsMap$Latitude, lon=MyZoopsMap$Longitude) 

station_zoop_points <- st_as_sf(MyZoopsMap, coords = c("Longitude", "Latitude"), 
                 crs= 4326, agr = "constant") %>% 
                  st_transform(crs=4326) %>% 
  mutate(ID=1:15) %>% 
  filter(ID>4) %>% 
  filter(ID != 11) %>% 
  filter(ID != 13) %>% 
  filter(ID != 7) %>% 
  mutate(ID=1:8)

# final list of stations
zoop_station_list <- station_zoop_points %>% 
  select(Station)
zoop_station_list

# final list of stations that match salinity table
station_zoop6 <- tibble(Station=MyZoopsMap$Station, lat=MyZoopsMap$Latitude, lon=MyZoopsMap$Longitude) %>% 
  filter(Station== "NZ028" | Station== "NZ048"| Station== "NZ054" | Station== "NZ060"|
           Station=="NZ064" | Station == "NZD41")

station_zoop_points6 <- station_zoop_points %>% 
  filter(Station== "NZ028" | Station== "NZ048"| Station== "NZ054" | Station== "NZ060"|
           Station=="NZ064" | Station == "NZD41")

# upstream most station
station_zoop_ref <- station_zoop_points %>% 
  filter(Station=="NZ064")

# clip stations to waterways
#station_zoop_clip <- st_intersection(ww_clip, station_zoop_points)

#Plot the spatial extent
plot <- ggplot()+
  geom_sf(data=Regions_mismatches)+
  geom_sf(data=station_zoop_points, col="red")+
  #scale_fill_grey(name="Region", start = 0.4)+
  #scale_color_viridis_d(name="Survey")+
  #scale_shape_manual(name="Survey", values=c(1,2,15,18,4))+
  labs(title="Mismatches study extent", subtitle="1980 to present")+
  theme_bw()
plot

# add all waterways
ggplot()+
  geom_sf(data=Regions_mismatches,color="gray90", alpha=0.5) +
  #geom_sf(data=waterways2, color="slategray1", fill="slategray1") +
  geom_sf(data=station_zoop_points, col="red")+
  theme_bw()

# add clipped waterways, with labels
ggplot()+
  #geom_sf(data=Regions_mismatches,color="gray90", alpha=0.5) +
  #geom_sf(data=ww_clip, color="slategray1", fill="slategray1") +
  geom_sf(data=station_zoop_points, col="red")+
  theme_bw()+
  geom_sf_label(data=station_zoop_points, aes(label=Station))

# calculate Euclidean distances between 8 points (no bounding) in units of meters (m) then transform
zoop_dist <- sf::st_distance(station_zoop_points, station_zoop_points, which = "Euclidean")
zoop_dist

# all 8 points
Station <- c("NZD16", "NZ064", "NZ060", "NZ054", "NZ048", "NZ028", "NZD06", "NZD41")
Dist <- c(0, 6400.284, 14659.032, 21748.739, 29107.359, 33370.744, 39826.917, 62091.305)
zoop_dist_df <- tibble(Station, Dist)

# calculate Euclidean distances between 5 points and one station further upstream
zoop_dist_up <- sf::st_distance(station_zoop_points6, station_zoop_ref) # meters only, transform later with vegan
zoop_dist_up
# Units: [m]
#           [,1]
# [1,]     0.000
# [2,]  8264.514
# [3,] 15358.062
# [4,] 22731.765
# [5,] 27293.046
# [6,] 55728.566

# reorder the 6 stations with matching salinity values - according to zoop_env and zoop_spp
Station <- c("NZD41", "NZ028", "NZ048", "NZ054", "NZ060", "NZ064")
Dist <- c(55728.566, 27293.046, 22731.765, 15358.062, 8264.514, 0.000)
zoop_dist_df <- tibble(Station, Dist)


ggplot()+
  geom_sf(data=Regions_mismatches,color="gray90", alpha=0.5) +
  #geom_sf(data=ww_clip, color="slategray1", fill="slategray1") +
  geom_sf(data=station_zoop_points6, col="red")+
  theme_bw()+
  geom_sf_label(data=station_zoop_points6, aes(label=Station))+
  labs(title="Zooplankton and salinity stations")


## for later if needed - calculate distances between points within the waterway polygon
# library(sfnetworks)
# https://stackoverflow.com/questions/68596244/can-i-bound-an-st-distance-call-by-a-polygon
# https://luukvdmeer.github.io/sfnetworks/index.html
# station_zoop_net <- sfnetwork(station_zoop_points, crs=4326)
# station_zoop_net
```

## 2 - zoop salinities
```{r}
MyZoopsMeso <- MyZoopsF %>% 
  filter(SizeClass == "Meso")
View(MyZoopsMeso)

# Filter species of interest
MyZoopsMeso_SppSal <- MyZoopsMeso %>% 
  filter(Species == "Eurytemora affinis" |
          Species == "Pseudodiaptomus forbesi" |
           Species=="Sinocalanus doerrii" |
           Species=="Limnoithona tetraspina" |
           Species=="Bosmina longirostris") %>%
  group_by(Station, Month, Year, Species, Salinity_zone) %>% 
  summarize(meanCPUE=mean(CPUE, na.rm=T), meanSalSurf=mean(SalSurf, na.rm=T)) %>% 
  right_join(zoop_station_list, by="Station") %>% 
  ungroup() %>%   
  filter(Year>1994)

View(MyZoopsMeso_SppSal)

check <- MyZoopsMeso_SppSal %>%
  ungroup() %>% 
  distinct(Station)
```

```{r}
MyZoopsSal <- MyZoopsMeso_SppSal %>% 
  select(Station, Month, Year, meanSalSurf) %>% 
  group_by(Station, Month, Year) %>% 
  summarize(meanSalSurf=mean(meanSalSurf, na.rm=T)) %>% 
  mutate(Day=1) %>% 
  ungroup()

check2 <- MyZoopsSal %>%
  ungroup() %>%
  distinct(Station)

MyZoopsSal <- unite(MyZoopsSal, "Date", c("Month","Day","Year"), sep="/", remove = FALSE)
MyZoopsSal$Date <- mdy(MyZoopsSal$Date)

# For imputation - Long to wide format
MyZoopsSal2 <- MyZoopsSal %>% 
  select(Date, Station, meanSalSurf)%>% 
  pivot_wider(id_cols = Date, names_from=Station, values_from=meanSalSurf, values_fill=NA) %>% 
  arrange(Date)
View(MyZoopsSal2)
```

```{r}
# create a template of dates
Date <- tibble(Date=seq(from=lubridate::as_date("1995-01-01"), 
                        to=lubridate::as_date("2020-12-01"), by="1 month")) # 312*8 = 2496

# create a new table with all combos
sal_table <- Date %>% 
  left_join(MyZoopsSal2, by="Date")
```


### arima imputation
```{r, eval=FALSE}
library(tidyverse)
library(ggthemes)
library(imputeTS)
library(zoo)
library(forecast)
library(patchwork)
library(sf)
```

Full model
For-loop and model output
```{r, eval=FALSE}
# prep dataframe for for-loop
salwodate <- sal_table %>% 
  arrange(Date) #%>%  # ensure it's in chronological order
salwodate <- salwodate[,-1] # then remove date and analyte columns
salwodate <- log10(salwodate + 1) # add one and log10 transform
summary(salwodate)
head(salwodate) # 6 cols


### Run this code altogether 
#create an empty matrix
Time_series_inter <- matrix(NA,nrow(salwodate),(ncol(salwodate))) #create empty matrix
dim(Time_series_inter)
colnames(Time_series_inter) <- colnames(salwodate)
head(Time_series_inter)

# Create a vector of years (for each month x year combo = 312)
Year <- lubridate::year(sal_table$Date)

## start for-loop, make sure to refresh Time_series_inter above
for (i in 1:ncol(salwodate)) {
  #create time-series object
  dat_imp <- data.frame(Year, y = salwodate[,i])

  y <- ts(dat_imp[2], start= c(1995,1), end = c(2020,12), frequency = 12) # time series object
  
  #add raw values into empty matrix
  Time_series_inter[,i] <- dat_imp[,c(-1)] # removing the Year column c(-1)
  
  #interpolate missing values (if any)
    if(length(which(is.na(y))) > 0){
          
        #fit ARIMA and impute missing values
        fit <- auto.arima(y, seasonal = TRUE) # don't use lambda, already log10 transformed 
        y_inter <- na_kalman(y,model=fit$model)

        #identify missing values to impute (and replace in the matrix)
        id.na <- which(is.na(y))

        #remove missing values at the begining and end of time series
        aa <- split(id.na, cumsum(c(1, diff(id.na) != 1)))  #split sequences of missing values 
        last <- length(y)
        first <- 1
        is.first <- sapply(aa,function(x) length(which(x %in% first))) #identify series of NAs at the start of the time-series
        is.last <- sapply(aa,function(x) length(which(x %in% last))) #idem with last value
          
        if(sum(is.first) > 0 | sum(is.last) > 0){
            aa <- aa[-c(which(is.last == 1),which(is.first == 1))] #remove them from the list
            id.na <- unlist(aa)
        }

        #replace missing values by imputed values into matrix
        Time_series_inter[id.na,i] <- y_inter[id.na]

# jpeg(paste0("Imputation_plots/salorophyll",colnames(Time_series_inter)[i],".jpeg"))
# plot(y_inter,xlab="",ylab="salorophyll (C)",col="red",main=(Time_series_inter)[i])
# lines(y,col="black")
# mtext(side=3,text=paste("Station #",colnames(Time_series_inter)[i]),line=-1,adj=0.9)
# dev.off()
# 
# 		}else{
# 		#save plot for non-imputed time-series
# 		jpeg(paste0("Imputation_plots/salorophyllRaw",colnames(Time_series_inter)[i],".jpeg"))
# 		plot(y,xlab="",ylab="salorophyll (C)",col="black",main=(Time_series_inter)[i])
# 		mtext(side=3,text=paste("Station #",colnames(Time_series_inter)[i]),line=-1,adj=0.9)
# 		dev.off()
		}
    print(i)
     
}
## end for-loop

#output of imputation
View(Time_series_inter)

# back transform
Time_series_inter2 <- 10^Time_series_inter -1

# evaluate top model
fit
```

Create new time series with interpolated values
```{r, eval=FALSE}
#show wide table of imputed values with dates and analyte type
Time_series_imp_wide <- as_tibble(Time_series_inter2) %>% 
  mutate(Date = sal_table$Date, Analyte= "Salinity") %>% # add columns back
  select(Date, Analyte, NZ028:NZD41)  # rearrange columns

Time_series_imp <- Time_series_imp_wide %>%
  pivot_longer(
    cols= "NZ028":"NZD41",
    names_to="Station", 
    values_to = "Imputed_values")
View(Time_series_imp)
summary(Time_series_imp)
dim(Time_series_imp) # 1872 rows - only 49 NAs remaining

# identify which values were imputed - which values DONT they have in common?
# If Imputed_values has a value, where the raw data has an NA, and Missing=TRUE, it's imputed
sal_table_long <- sal_table %>% 
  pivot_longer(
    cols="NZ028":"NZD41",
    names_to="Station",
    values_to = "Raw_values"
  ) %>% 
  mutate(Missing=ifelse(is.na(Raw_values), TRUE, FALSE)) # Is it missing?
View(sal_table_long)
summary(sal_table_long)
dim(sal_table_long) # 1872 rows- with 108 NAs present

# Join them to create a master table tracking which values were modeled
# Should still be 10 NAs that are missing but not imputed
Time_series_imp2 <- Time_series_imp %>% 
  left_join(sal_table_long) %>% 
  mutate(Imputed=ifelse(Missing==TRUE & Imputed_values>0, TRUE, FALSE)) 
View(Time_series_imp2)
```

Diagnostic tests
```{r, eval=FALSE}
#Plot the observed and predicted salinity
ggplot(Time_series_imp2, aes(Date, Imputed_values, color=Imputed))+
  geom_line(aes(Date, Imputed_values), col="gray60")+  
  geom_point(size=1)+
  facet_wrap(Station~., ncol=3, scales = "free_y")+
  scale_color_manual(values=c("gray60", "red", "black"))+
  labs(title="Replacing missing values with ARIMA and Kalman filter", subtitle="Salinity")+
  theme_bw()
```

```{r, eval=FALSE}
# model summary
summary(fit)

# Check the residuals
forecast::checkresiduals(fit)

# Check several metrics of performance
forecast::accuracy(object = fit)

#plot residuals versus fitted values (=check for heterosedasticity)
#if problems you might want to try to transform the data first
par(mar=c(4,4,4,4))
plot(fit$fitted, fit$residuals,ylab="Residuals",xlab="Fitted values")
```

Export df
```{r, eval=FALSE}
Time_series_imp2 %>% write_csv("mismatches_salinity_imputed_data.csv")
```

Re-import df
```{r, eval=TRUE}
Time_series_imp2 <- read_csv("mismatches_salinity_imputed_data.csv")
```

Make wide dataframe
```{r, eval=TRUE}
Time_series_imp3 <- Time_series_imp2 %>% 
  select(Date, Station, Imputed_values) %>% 
  filter(!is.na(Imputed_values))
View(Time_series_imp3)
```


## 3 - zoop abundances
```{r}
MyZoopsMeso_Spp <- MyZoopsMeso_SppSal %>% 
  mutate(Day=1)
MyZoopsMeso_Spp <- unite(MyZoopsMeso_Spp, "Date", c("Month","Day","Year"), sep="/", remove = FALSE)
MyZoopsMeso_Spp$Date <- lubridate::mdy(MyZoopsMeso_Spp$Date)

MyZoopsMeso_Spp_w <- MyZoopsMeso_Spp %>% 
  select(Date, Station, Species, meanCPUE) %>% 
  pivot_wider(id_cols=c("Date", "Station"), names_from="Species", values_from="meanCPUE")

# find the common samples - both salinity and zoops
MyZoopsCompleteSamples <- MyZoopsMeso_Spp_w %>% 
  inner_join(Time_series_imp3, by=c("Date", "Station")) %>% 
  rename(salimp=Imputed_values) %>% 
  arrange(Date,Station) %>% 
  #left_join(zoop_dist_df, by="Station") # use this for precalculated distances
  left_join(station_zoop6, by="Station") %>% 
  mutate(Total=`Bosmina longirostris` + `Eurytemora affinis` + `Limnoithona tetraspina` +
                     `Pseudodiaptomus forbesi` + `Sinocalanus doerrii`) %>% 
  filter(Total>0)
View(MyZoopsCompleteSamples)

# for vegan package
zoop_env <- MyZoopsCompleteSamples %>% 
  select(Station, Date, salimp)
zoop_env <- unite(zoop_env, "Station_date", c("Station", "Date"), remove=TRUE)
zoop_env <- zoop_env %>% 
  column_to_rownames("Station_date")
View(zoop_env)
summary(zoop_env)

# for vegan package
zoop_spp <- MyZoopsCompleteSamples %>% 
  select(-salimp, -lat, -lon)
zoop_spp <- unite(zoop_spp, "Station_date", c("Station", "Date"), remove=TRUE)
zoop_spp <- zoop_spp %>% 
  column_to_rownames("Station_date") 
View(zoop_spp)
summary(zoop_spp)

# for vegan package
zoop_latlon <- MyZoopsCompleteSamples %>%
  #select(Station, Date, Dist) %>% #for precalculated distances
  select(Station, Date, lat, lon)
zoop_latlon <- unite(zoop_latlon, "Station_date", c("Station", "Date"), remove=TRUE)
zoop_latlon <- zoop_latlon %>% 
  column_to_rownames("Station_date")
View(zoop_latlon)
summary(zoop_latlon)

# for vegan package
zoop_date <- MyZoopsCompleteSamples %>%
  select(Station,Date)
zoop_date <- unite(zoop_date, "Station_date", c("Station", "Date"), remove=FALSE)
zoop_date <- zoop_date %>% 
  select(Station_date, Date) %>% 
  column_to_rownames("Station_date")
View(zoop_date)
summary(zoop_date)
```

## 4 - zoop varpart

## load packages
```{r}
#install.packages ('vegan')
library(vegan)
```


Tutorial URL: https://www.davidzeleny.net/anadat-r/doku.php/en:varpart_examples

*Functions*
varpart (library vegan) - variation partitioning (using RDA, CCA or db-RDA) among up to four matrices of environmental variables. The first argument (Y) is the dependent variable, usually the matrix of species composition (the function calculates RDA, or, if chisquare = TRUE, CCA), but could be also only a single variable (in that case it calculates linear regression) or distance matrix (applying db-RDA using the function capscale). Further arguments (up to four) are (groups of) explanatory variables. The function uses either formula interface (with ~, see examples) or matrices. The interpretation should be based on adjusted R2, although raw R2 is also reported (for CCA, adjusted R2 is calculated by permutation method of Peres-Neto et al. 2006 and may slightly vary between re-analyses of the same data; the argument permutations specifies the number of permutations used to calculate adjusted R2 in CCA).

plot.varpart (library vegan) - draws Venn's diagram with fractions of explained variation. In default setting it doesn't show negative values of explained variation (argument cutoff = 0). The function can use arguments of showvarparts below, e.g. to add the labels for individual (groups of) variables (argument Xnames), or colours of the fractions (argument bg). Consult ?plot.varpart for more details.

showvarparts (library vegan) - draws schema of Venn's diagram with codes of individual fractions.

*Description of the dataset*
The experiment was focused on diversity and species richness of weed communities in barley fields after application of industrial fertilizers. The experiment was originally established as a part of hydrobiological research, and for practical reasons it was not established using properly randomized design - individual plots are practically pseudoreplications, but for the purpose of this exercise we ignore it.

All together 122 plots of the size 1 m2 were fertilized by three different fertilizers (for the purpose of this study the types of fertilizers are not distinguished), and on each plot species composition of weed community was recorded (with species covers estimated using ordinal scale from 1 to 7). Fertilizing simultaneously influences also growth and cover of barley, and the cover of barley can in turn also influence species composition of weeds; for this reason, the barley cover in the plot was also recorded.

## import data
```{r}
fertil.spe <- read.delim ('https://raw.githubusercontent.com/zdealveindy/anadat-r/master/data/fertil.spe.txt', row.names = 1)
fertil.env <- read.delim ('https://raw.githubusercontent.com/zdealveindy/anadat-r/master/data/fertil.env.txt', row.names = 1)
```

Species data in fertil.spe are represented by the estimated cover of weed community species in a plot, and it is not necessary to transform them. One may first apply DCA ordination on species data to see the length of the first DCA axis and hence seek the recommendation whether the linear or unimodal method should be used (the length is 3.8 S.D., lying in a grey zone between 3 and 4 S.D.); we will use Redundancy analysis (RDA) here.

First, let's see how would the variation partitioning looks like if we do it step-by-step manually (without using varpart function). We will partition variation explained by variables dose and cover from fertil.env. For this, we need to define the global model (to see how much variation is explained by both of them), and also models with simple (marginal) effect of each explanatory variable:

## model 1 - manual
```{r}
# fractions [a+b+c]:
rda.all <- rda (fertil.spe ~ dose + cover, data = fertil.env)
# fractions [a+b]:
rda.dose <- rda (fertil.spe ~ dose, data = fertil.env)
# fractions [b+c]:
rda.cover <- rda (fertil.spe ~ cover, data = fertil.env)
```

Now we can calculate variations explained by individual fractions (using RsquareAdj function):
```{r}
## fraction [a+b+c]
RsquareAdj (rda.all)
# $`r.squared`
# [1] 0.1430381
# 
# $adj.r.squared
# [1] 0.1286354
abc <- RsquareAdj (rda.all)$adj.r.squared
 
## fraction [a+b]:
RsquareAdj (rda.dose)
# $`r.squared`
# [1] 0.06889364
# 
# $adj.r.squared
# [1] 0.06113442
ab <- RsquareAdj (rda.dose)$adj.r.squared
 
## fraction [b+c]:
RsquareAdj (rda.cover)
# $`r.squared`
# [1] 0.09628412
# 
# $adj.r.squared
# [1] 0.08875315
bc <- RsquareAdj (rda.cover)$adj.r.squared
```

In this way, we calculated adjusted R2 of fractions [a+b+c], [a+b] and [b+c]. To calculate values for individual fractions [a], [b] and [c], we need to do a simple subtractions:

```{r}
b <- ab + bc - abc  # 0.02125216
a <- ab - b         # 0.03988225
c <- bc - b         # 0.06750099
```


After recalculating into percentage, we get:

[a] = 4.0%
[b] = 2.1%
[c] = 6.8%
Conditional effect of dose (fraction [a] = 4.0%) is slightly lower than conditional effect of cover ([c] = 6.8%), and shared variance is not too high ([b] = 2.1%). Seems that weed species composition is more affected by light conditions modified by increasing cover of barley, than by fertilizer itself.

Now, let's see the same, using function varpart (note that by default, varpart function calculates RDA on the matrices; if you want to use CCA, you need to use argument chisquare = TRUE).

## model 2 - varpart
```{r}
varp <- varpart (fertil.spe, ~ dose, ~ cover, data = fertil.env)
varp
```

```{r}
varp_zoop2 <- varpart(vegdist(zoop_spp, method="bray"), zoop_env, zoop_latlon)
varp_zoop2

varp_zoop3 <- varpart(vegdist(zoop_spp, method="bray"), zoop_env, zoop_latlon, zoop_date)
varp_zoop3
```
Partition of squared Bray distance in dbRDA 

Call: varpart(Y = vegdist(zoop_spp), X = zoop_env, zoop_eucdist)

Explanatory tables:
X1:  zoop_env
X2:  zoop_eucdist 

No. of explanatory tables: 2 
Total variation (SS): 354.84 
No. of observations: 1764 

Partition table:
                     Df R.squared Adj.R.squared Testable
[a+b] = X1            1   0.16939       0.16892     TRUE
[b+c] = X2            1   0.28182       0.28142     TRUE
[a+b+c] = X1+X2       2   0.33368       0.33292     TRUE
Individual fractions                                    
[a] = X1|X2           1                 0.05150     TRUE
[b]                   0                 0.11742    FALSE
[c] = X2|X1           1                 0.16400     TRUE
[d] = Residuals                         0.66708    FALSE

Use function dbrda to test significance of fractions of interest

```{r}
#Plot all the variables of interest, do any variables seem highly related to
#the geographic variables?
rda.both <- rda(zoop_spp, zoop_env, zoop_latlon)
plot(rda.both, choices=c(1,2), type="text")

#Let's see how our just environmental data looks
rda.env <- rda(zoop_spp ~., data=zoop_env)
plot(rda.env, choices=c(1,2), type="text")

#Let's also see how things look just based on latitude and longitude.
#Does it look like some variance is explained just by geographical gradients?
rda.geo <- rda(zoop_spp ~ ., data=zoop_latlon)
plot(rda.geo, choices=c(1,2), type="text")
```


Partition of variance in RDA 

Call: varpart(Y = fertil.spe, X = ~dose, ~cover, data = fertil.env)

Explanatory tables:
X1:  ~dose
X2:  ~cover 

No. of explanatory tables: 2 
Total variation (SS): 2581.2 
            Variance: 21.332 
No. of observations: 122 

Partition table:
                     Df R.squared Adj.R.squared Testable
[a+b] = X1            1   0.06889       0.06113     TRUE
[b+c] = X2            1   0.09628       0.08875     TRUE
[a+b+c] = X1+X2       2   0.14304       0.12864     TRUE
Individual fractions                                    
[a] = X1|X2           1                 0.03988     TRUE
[b]                   0                 0.02125    FALSE
[c] = X2|X1           1                 0.06750     TRUE
[d] = Residuals                         0.87136    FALSE

Use function rda to test significance of fractions of interest

The results for individual fractions are identical with our results above. varpart reports also simple (marginal) effect (both not-adjusted and adjusted variation) of individual predictors, i.e. fractions [a+b] and [b+c].

We may plot the results into Venn's diagram (argument digits influences number of decimal digits shown in the diagram, Xnames the displayed names of variables, and bg background color of the diagram fractions; see ?varpart for details):

## plot venn-diagram
```{r}
plot (varp, digits = 2, Xnames = c('dose', 'cover'), bg = c('navy', 'tomato'))
```

```{r}
plot (varp_zoop2, digits = 2, Xnames = c('salinity', 'lat/lon'), bg = c('navy', 'tomato'))
plot (varp_zoop3, digits = 2, Xnames = c('salinity', 'lat/lon', 'date'), bg = c('navy', 'tomato', 'gray'))
```


Now, when we know both simple and conditional effect of each variables, we may want to know whether these variances are significant, and hence worth of interpreting. Results from varpart contain the column testable with logical values indicating whether given fraction is testable or not. To test each of them, we will need the models defined above, and the function anova, which (if applied on single object resulting from rda or cca method, returns Monte Carlo permutation test of the predictor effect). For this, we need to first define also partial ordination models with one variable as exlanatory and the other as covariable (Condition):

```{r}
# fraction [a]:
rda.dose.cover <- rda (fertil.spe ~ dose + Condition (cover), data = fertil.env)
# fraction [c]:
rda.cover.dose <- rda (fertil.spe ~ cover + Condition (dose), data = fertil.env)
```

```{r}
# for varp_zoop2

# fraction [a]
rda.sal.dist <- rda(zoop_spp ~ zoop_env$salimp + Condition(zoop_latlon$lat + zoop_latlon$lon))
rda.sal.dist
# fraction [c]
rda.dist.sal <- rda(zoop_spp ~ zoop_latlon$lat + zoop_latlon$lon + Condition(zoop_env$salimp))
rda.dist.sal


## The global model (fractions [a+b+c]): # sig - 0.006
anova(rda.both)

## Simple (marginal) effect of salimp (fraction [a+b]): # sig - 0.005
anova(rda.env)

## ## Simple (marginal) effect of Dist (fraction [b+c]): # sig - 0.04
anova(rda.geo)

## Conditional (partial) effect of salimp (fraction [a]): # sig - 0.01
anova(rda.sal.dist)

## Conditional (partial) effect of Dist (fraction [c]): # not sig
anova(rda.dist.sal)
```


Then, to test individual fractions, we use:

```{r}
## The global model (fractions [a+b+c]):
anova (rda.all)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = fertil.spe ~ dose + cover, data = fertil.env)
# Df Variance      F Pr(>F)    
# Model      2   3.0513 9.9313  0.001 ***
#   Residual 119  18.2808                  
# ---
#   Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1
 
## Simple (marginal) effect of dose (fraction [a+b]):
anova (rda.dose)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = fertil.spe ~ dose, data = fertil.env)
# Df Variance      F Pr(>F)    
# Model      1   1.4696 8.8789  0.001 ***
#   Residual 120  19.8624                  
# ---
#   Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1
 
## ## Simple (marginal) effect of cover (fraction [b+c]):
anova (rda.cover)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = fertil.spe ~ cover, data = fertil.env)
# Df Variance      F Pr(>F)    
# Model      1   2.0539 12.785  0.001 ***
#   Residual 120  19.2781                  
# ---
#   Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1
 
## Conditional (partial) effect of dose (fraction [a]):
anova (rda.dose.cover)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = fertil.spe ~ dose + Condition(cover), data = fertil.env)
# Df Variance      F Pr(>F)    
# Model      1   0.9974 6.4924  0.001 ***
#   Residual 119  18.2808                  
# ---
#   Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1
 
## Conditional (partial) effect of cover (fraction [c]):
anova (rda.cover.dose)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = fertil.spe ~ cover + Condition(dose), data = fertil.env)
# Df Variance      F Pr(>F)    
# Model      1   1.5817 10.296  0.001 ***
#   Residual 119  18.2808                  
# ---
#   Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1
```

From these results, you may see that all simple (marginal) and conditional (partial) effects of both predictors are significant at P < 0.001.


# Fish varpart
```{r}
# MWT data
Fish_SppMonthStn <- BayStudyFA %>% 
  group_by(Station, Date, Taxa) %>% 
  summarize(meanCPUE=mean(Count, na.rm=T)) %>% 
  pivot_wider(id_cols=c("Station", "Date"), names_from="Taxa", values_from="meanCPUE", values_fill = 0) %>% 
  mutate(Total= `Engraulis mordax` + `Clupea pallasii` + `Spirinchus thaleichthys` + `Morone saxatilis` +
           `Dorosoma petenense` + `Alosa sapidissima` + `Hypomesus transpacificus`) %>% 
  filter(Total>0) %>% 
  select(-Total)
View(Fish_SppMonthStn)
summary(Fish_SppMonthStn)

Fish_stn_date <- Fish_SppMonthStn %>% 
  select(Station, Date)

Fish_sal <- BayStudyFA %>% 
  group_by(Station, Date) %>% 
  summarize(meanSal = mean(Sal_surf, na.rm=TRUE)) %>% 
  right_join(Fish_stn_date, by = c("Station", "Date"))
summary(Fish_sal)

Fish_geo <- BayStudyFA %>% 
  select(Station) %>% 
  distinct() %>% 
  left_join(BayStudyMap, by="Station") %>% 
  select(Station, Latitude, Longitude)

Fish_CompleteSamples <- Fish_SppMonthStn %>% 
  left_join(Fish_sal, by = c("Station", "Date")) %>% 
  left_join(Fish_latlon, by="Station")
summary(Fish_CompleteSamples)

# for vegan
fish_spp <- Fish_SppMonthStn
fish_spp <- unite(fish_spp, "Station_date", c("Station", "Date"), remove=TRUE)
fish_spp <- fish_spp%>% 
  column_to_rownames("Station_date")
dim(fish_spp)

fish_env <- Fish_sal
fish_env <- unite(fish_env, "Station_date", c("Station", "Date"), remove=TRUE)
fish_env <- fish_env %>% 
  column_to_rownames("Station_date")
dim(fish_env)

fish_latlon <- Fish_CompleteSamples %>% 
  select(Station, Date, Latitude, Longitude)
fish_latlon <- unite(fish_latlon, "Station_date", c("Station", "Date"), remove=TRUE)
fish_latlon <- fish_latlon %>% 
  column_to_rownames("Station_date")
dim(fish_latlon)
```

```{r}
fish_varpart <- varpart(vegdist(fish_spp, method="bray"), fish_env, fish_latlon)
fish_varpart

plot (fish_varpart, digits = 2, Xnames = c('salinity', 'lat/lon'), bg = c('navy', 'tomato'))
```


# Peak timing: peak, range, CV on an annual basis
- Tidyverse Slicemax function to ID peaks
- Plot distributions of CVs of peak timing - fish vs. zoop
- Regressions between peak timing and environmental variables (depth, salinity, temperature, secchi, flow)

## Zoop
```{r}

```

## Fish
```{r}

```

